Confidence Intervals for Prevalence Estimates from Complex Surveys with Imperfect Assays

### 2021-11-24 PrevAdjSurvey.tex
Updates from Overleaf

where \( w_i = N_i / \sum_{j=1}^n N_j \) and, therefore, \( \sum_{i=1}^k w_i = 1 \).

@@

\begin{equation}
    \hat{\beta}^* = \frac{\sum_{i=1}^k w_i \hat{\theta}_i}{\hat{\phi}_p - \hat{\phi}_n} - \frac{\hat{\phi}_n}{\hat{\phi}_p - \hat{\phi}_n}.
\end{equation}

@@

First, we consider the scenario0 \( k = 1 \), \( \phi_n > 0 \), and \( \phi_p < 1 \).
We develop a confidence interval for the population prevalence, \( \beta^* \).
Because \( k = 1 \), the estimator in Equation~\ref{eq:long-beta} becomes

\begin{equation}
\hat{\beta}^* = \frac{\hat{\theta}_1 - \hat{\phi}_n}{\hat{\phi}_p - \hat{\phi}_n}.
\label{eq:srs-beta-est}
\end{equation}

To create a confidence interval for \( \hat{\beta}^* \), we use a generalization of the melding method \cite{FayP:2015}, which makes use of lower and upper confidence distributions on functions of independent estimators to account for variability in \( \hat{\theta}_1 \), \( \hat{\phi}_n \), and \( \hat{\phi}_p \).

Each estimated component in Equation~\ref{eq:srs-beta-est} is a binomial probability parameter.
For each of these, we use distributions associated with  the exact binomial confidence interval.
For a binomial experiment with \( x \) successes out of \( n \) trials, the lower confidence distribution \( (B^L) \) is \( \operatorname{Beta}(x, n - x + 1) \), and the upper confidence distribution \( (B^U) \) is \( \operatorname{Beta}(x + 1, n - x)\).

Let \( q(a, W) \) be the \( a \)th quantile of a random variable \( W \).

Then the \( a \)\% confidence interval for \( \beta^* \) is

\begin{equation}
    \left\{ q \left( \frac{1 - a}{2}, \frac{B_{\theta_1}^L - B_{\phi_n}^L }{B_{\phi_p}^U - B_{\phi_n}^L }  \right),  q \left( \frac{1 + a}{2}, \frac{B_{\theta_1}^U - B_{\phi_n}^U}{B_{\phi_p}^L - B_{\phi_n}^U}  \right) \right\}.
\label{eq:srs-conf-int}
\end{equation}

@@

% \begin{figure}
% \centering
% \includegraphics[width=0.8\textwidth]{figures/simple_coverage_error_comparison_plot.pdf}
% \caption{Caption}
% \label{fig:simple_coverage_error_comparison_plot}
% \end{figure}

>>>>

where \( w_i = N_i / \sum_{j=1}^k N_j \) and, therefore, \( \sum_{i=1}^k w_i = 1 \).

@@

\begin{equation}
    \hat{\beta}^* = \frac{\sum_{i=1}^k w_i \hat{\theta}_i}{\hat{\phi}_p - \hat{\phi}_n} - \frac{\hat{\phi}_n}{\hat{\phi}_p - \hat{\phi}_n}. \label{eq:betastarhat}
\end{equation}

@@

First, we consider the scenario0 \( k = 1 \), \( \phi_n > 0 \), and \( \phi_p < 1 \).
We develop a confidence interval for the population prevalence, \( \beta^* \).
When \( k = 1 \), the estimand in Equation~\ref{eq:long-beta} becomes $\beta^* = (\theta_1 - \phi_n)/(\phi_p-\phi_n)$. We have $\phi_p > \phi_n$ for any useful assay, and when the sample is a mixture of individuals with and without the disease of interest, $\phi_p > \theta_1 > \phi_n$. The estimator of $\beta^*$ is 
%in Equation~\ref{eq:betastarhat} becomes

\begin{equation}
\hat{\beta}^* \equiv 
g(\hat{\theta}_1, \hat{\phi}_n, \hat{\phi}_p)
\equiv 
\left\{ 
\begin{array}{ll}
1 & \mbox{ if $\hat{\phi}_n < \hat{\phi}_p \leq \hat{\theta}_1$ }  \\
\frac{\hat{\theta}_1 - \hat{\phi}_n}{\hat{\phi}_p - \hat{\phi}_n} & 
\mbox{ if $\hat{\phi}_p > \hat{\theta}_1 > \hat{\phi}_n$ } \\
0 & \mbox{ otherwise} 
\end{array}
\right.
\label{eq:srs-beta-est}
\end{equation}
where we define $0/0=0$.
% Damon, that definition is arbitrary, it is just used so that everything is defined

To create a confidence interval for \( \hat{\beta}^* \), we use a generalization of the melding method \cite{FayP:2015}, which makes use of lower and upper confidence distributions on functions of independent estimators to account for variability in \( \hat{\theta}_1 \), \( \hat{\phi}_n \), and \( \hat{\phi}_p \). Confidence distributions are like frequentist posterior distributions
\cite{}.  The lower and upper confidence distributions are used with discrete responses to ensure the validity of the resulting inferences, and for the binomial case they are equivalent to the posterior distributions that result from using well-calibrated null preference priors \cite{}.

Each estimated component in Equation~\ref{eq:srs-beta-est} is a binomial probability parameter.
For each of these, we use distributions associated with  the exact binomial confidence interval.
For a binomial experiment with \( x \) successes out of \( n \) trials, the lower confidence distribution \( (B^L) \) is \( \operatorname{Beta}(x, n - x + 1) \), and the upper confidence distribution \( (B^U) \) is \( \operatorname{Beta}(x + 1, n - x)\).
Let \( q(a, W) \) be the \( a \)th quantile of a random variable \( W \). Then the exact \( a \)\% central confidence interval of Clopper-Pearson for the binomial parameter is 
\[
\left\{ q \left( \frac{1-a}{2}, B^L \right), q \left( \frac{1+a}{2}, B^U \right\}.
\]

Fay et al \cite{FayP:2015} proposed a method for obtaining confidence intervals for functions of two parameters that are monotonic within the allowable range  for each parameter given the other is fixed. Here we generalize that to $\beta^*$, which is a function of 3 parameters. When $1 \geq \phi_p > \theta_1 > \phi_n \geq 0$ then $\beta^*$ is monotonically increasing in $\theta_1$, monotonically decreasing in $\phi_p$, and monotonically decreasing in $\phi_n$.
Then the \( a \)\% confidence interval for \( \beta^* \) is

\begin{equation}
    \left\{ q \left( \frac{1 - a}{2}, g \left\{ B_{\theta_1}^L, B_{\phi_n}^U, B_{\phi_p}^U \right\}   \right),  
            q \left( \frac{1 + a}{2},  g \left\{ B_{\theta_1}^U, B_{\phi_n}^L, B_{\phi_p}^L \right\}   \right) \right\}.
\label{eq:srs-conf-int}
\end{equation}

@@
